---
title: "Modeling"
author: "Lauren Bradley"
date: "12/1/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
# Load select_winsorized_df for analysis. 
# This dataframe includes variables that have more than 10,000 observations.
# This dataframe also replaced outliers with quantiles. 

load("C:/Users/laure/git/capstone/selected_Winsorized_df.RData")

```


First start with variable selection to find which variables appear the most. 

```{r}
library(tidyverse)
# On;y the quantitative variables. 
predictorSet <- c("LBDHDD", "LBDHDDSI", "LBXTR","LBDTRSI" , "LBDLDL"  , "LBDLDLSI", "LBXTC"   , "LBDTCSI" , "LBXGLU"  , "LBDGLUSI" ,"CBD121"  , "CBD131"  ,"DBD900"  , "DBD905"  , "DBD910"  , "PAQ625"  , "PAD630"  , "PAQ670"  , "PAD675"  , "PAD680"  , "WHD010"  , "WHD020"  , "WHD050"  , "WHD110", "WHD120" ,  "WHD130"   ,"WHD140" ,  "WHQ150" ,  "BMXWT"  ,  "BMXHT"  ,  "BMXBMI" ,  "BMXLEG" ,  "BMXARML"  ,"BMXARMC",  "BMXWAIST", "BPD035"  ,"INDFMPIR", "WTINT2YR", "WTMEC2YR")

include_df <- subset(selected_df, select = c(predictorSet))

#One of the big outliers of 9999 and 7777
include_df$WHD110[include_df$WHD110 == 9999 | include_df$WHD110 == 7777] <- NA
include_df$WHD120[include_df$WHD120 == 9999] <- NA

include_df$DIQ010 <- selected_df$DIQ010
include_df$Age <- selected_df$RIDAGEYR
include_df$Gender <- selected_df$RIAGENDR


# Including everything from include_df results in 0
include_df %>% 
  select(one_of(names(include_df))) %>% 
  na.omit() %>% 
  nrow()
include_df <- na.omit(include_df)

summary(include_df)
```
Include df has only quantitative values in it and the gender + age + diabetes variable. It also has 296 observations.


```{r}
mod1 <- glm(DIQ010 ~ LBDHDD
 + LBDHDDSI
 + LBXTR   
 + LBDTRSI 
 + LBDLDL  
 + LBDLDLSI
 + LBXTC
 + LBDTCSI
 + LBXGLU
 + LBDGLUSI
 + CBD121
 + CBD131
 + DBD900
 + DBD905
 + DBD910
 + PAQ625
 + PAD630
 + PAQ670
 + PAD675
 + PAD680
 + WHD010
 + WHD020
 + WHD050
 + WHD110
 + WHD120
 + WHD130 
 + WHD140 
 + WHQ150  
 + BMXWT  
 + BMXHT 
 + BMXBMI
 + BMXLEG 
 + BMXARML 
 + BMXARMC 
 + BMXWAIST
 + BPD035  
 + INDFMPIR
 + WTINT2YR
 + WTMEC2YR + Gender + Age,
 data = include_df, family = "binomial")
summary(mod1)

mod0 <- glm(DIQ010 ~ 1,data = include_df, family = "binomial")

model_step_AIC <- step(mod0,
                       scope = list(lower = mod0, upper = mod1),
                       direction="forward", k = 2, trace = 0)
summary(model_step_AIC)
```

LBXGLU - Fasting Glucose (mg/dL)
WHD140 - Self-reported greatest weight (pounds)
PAD675 - Minutes moderate recreational activities
LBXTC - Total Cholesterol (mg/dL)
BMXARML - Upper Arm Length (cm)
INDFMPIR - Ratio of family income to poverty
PAQ670 - Days moderate recreational activities
LBDHDDSI - Direct HDL-Cholesterol (mmol/L)
WTMEC2YR - Full Sample 2 Year MEC Exam Weight




Mod2 is a similar model to model_step_aic but without the insignificant variables

```{r}
mod2 <- glm(formula = DIQ010 ~ LBXGLU + WHD140 + PAD675 + LBXTC + BMXARML + PAQ670 + LBDHDDSI + WTMEC2YR, family = "binomial", 
    data = include_df)
summary(mod2)
```


```{r}
anova(mod2, model_step_AIC)
```

The anova says that mod 2 is better than the stepwise

```{r}
mod3 <- glm(formula = DIQ010 ~ LBXGLU + WHD140 + PAD675 + LBXTC + BMXARML + LBDHDDSI + WTMEC2YR, family = "binomial", 
    data = include_df)
summary(mod3)
```

```{r}
anova(mod3, model_step_AIC)
anova(mod3, mod2)
```

mod3 is better than the step wise and mod2 based on the anova. 



```{r}
mod_vars <- subset(include_df, select = c(
LBXGLU , WHD140 , PAD675 , LBXTC , BMXARML , INDFMPIR , 
    PAQ670 , LBDHDDSI , WTMEC2YR
))

pairs(mod_vars)
```

None of the variables look colinear except maybe WHD140 and BMXARML


Creating a tree to see other significant variables. 

```{r}
library(tree)
tree1 <- tree(DIQ010 ~ LBDHDD
 + LBDHDDSI
 + LBXTR   
 + LBDTRSI 
 + LBDLDL  
 + LBDLDLSI
 + LBXTC
 + LBDTCSI
 + LBXGLU
 + LBDGLUSI
 + CBD121
 + CBD131
 + DBD900
 + DBD905
 + DBD910
 + PAQ625
 + PAD630
 + PAQ670
 + PAD675
 + PAD680
 + WHD010
 + WHD020
 + WHD050
 + WHD110
 + WHD120
 + WHD130 
 + WHD140 
 + WHQ150  
 + BMXWT  
 + BMXHT 
 + BMXBMI
 + BMXLEG 
 + BMXARML 
 + BMXARMC 
 + BMXWAIST
 + BPD035  
 + INDFMPIR
 + WTINT2YR
 + WTMEC2YR + Gender + Age, data = include_df)

summary(tree1)
```

Tree 1 Variables Used:

LBXGLU - Fasting Glucose (mg/dL)
BMXWT - Weight (kg)
WTINT2YR - Full Sample 2 Year Interview Weight
PAQ625 - Number of days moderate work
WHD120 - Self-reported weight-age 25 (pounds)
BPD035 - Age told had hypertension
INDFMPIR - Ratio of family income to poverty
LBDHDD - Direct HDL Cholesterol (mg/dL)
LBXTR - Triglyceride (mg/dL)
WHD020 - Current self-reported weight (pounds)
WHD010 - Current self-reported height (inches)
PAD630 - Minutes moderate-intensity work
LBXTC - Total Cholesterol (mg/dL)
DBD910 - # of frozen meals/pizza in past 30 days



```{r}
plot(tree1)
text(tree1, cex = 0.7)
```


```{r}
tree_df <- subset(include_df, select = c(
  "LBXGLU" ,  "BMXWT"   , "WTINT2YR", "PAQ625"  , "WHD120",   "BPD035"  , "INDFMPIR", "LBDHDD"  , "LBXTR" ,   "WHD020"  , "WHD010"  , "PAD630",     
 "LBXTC"  ,  "DBD910"  
))
pairs(tree_df[1:5])
pairs(tree_df[5:10])
pairs(tree_df[10:14])
```

No extreme outliers so that is good.


------ MAIN VARIABLES I AM INTERESTED IN NOW -----------------------------

LBXGLU - Fasting Glucose (mg/dL)
BMXWT - Weight (kg)
WTINT2YR - Full Sample 2 Year Interview Weight
PAQ625 - Number of days moderate work
WHD120 - Self-reported weight-age 25 (pounds)
BPD035 - Age told had hypertension
INDFMPIR - Ratio of family income to poverty
LBDHDD - Direct HDL Cholesterol (mg/dL)
LBXTR - Triglyceride (mg/dL)
WHD020 - Current self-reported weight (pounds)
WHD010 - Current self-reported height (inches)
Age
LBXTC - Total Cholesterol (mg/dL)
DBD910 - # of frozen meals/pizza in past 30 days
PAD675 - Minutes moderate recreational activities
WTMEC2YR - Full Sample 2 Year MEC Exam Weight
PAQ670 - Days moderate recreational activities
WHD110 - Self-reported weight-10 yrs ago (pounds)
BMXBMI - Body Mass Index (kg/m**2)
LBDHDDSI - Direct HDL-Cholesterol (mmol/L)
BMXARML - Upper Arm Length (cm)
CBD121 - Money spent on eating out
PAD630 - Minutes moderate-intensity work
WHD140 - Self-reported greatest weight (pounds)


```{r}
# Selecting variables based on the ones listed. 

var_selected <- subset(selected_df, select = c(
LBXGLU ,
BMXWT ,
WTINT2YR ,
PAQ625 ,
WHD120 ,
BPD035 ,
INDFMPIR ,
LBDHDD ,
LBXTR ,
WHD020 ,
WHD010 ,
RIDAGEYR,
LBXTC ,
DBD910 ,
PAD675 ,
WTMEC2YR ,
PAQ670 ,
WHD110 ,
BMXBMI ,
LBDHDDSI ,
BMXARML ,
CBD121 ,
PAD630 ,
WHD140 ,
  DIQ010
))

head(var_selected)
```


Clean var_selected and delete rows with NA

```{r}
var_selected$WHD140[var_selected$WHD140 == 9999 | var_selected$WHD140 == 7777] <- NA

var_selected %>% 
  select(one_of(names(var_selected))) %>% 
  na.omit() %>% 
  nrow()
var_selected <- na.omit(var_selected)

var_selected$DIQ010 <- droplevels(var_selected$DIQ010)
```

Now with the selected variables, 515 observations are available. 


```{r}
null_mod <- glm(DIQ010 ~ 1, data = var_selected, family = "binomial")
full_mod <- glm(DIQ010 ~ ., data = var_selected, family = "binomial")
best_mod <- step(null_mod, scope = list(lower = null_mod, upper = full_mod),
                       direction="forward", k = 2, trace = 0)
summary(null_mod)
summary(full_mod)
summary(best_mod)
```



```{r}
tree2 <- tree(DIQ010 ~ LBXGLU + WTINT2YR + PAD675 + LBXTC + BMXBMI + 
    RIDAGEYR, data = var_selected)
summary(tree2)
```

```{r}
plot(tree2)
text(tree2, cex = 0.7)
```



```{r}
rforest2 <- randomForest(DIQ010 ~ ., data = var_selected, ntree = 250, nodesize = 25, importance = TRUE)
varImpPlot(rforest2, type = 1, main = "")
```

This plot shows the most important variables. 

```{r}
pairs(var_selected[1:7])
pairs(var_selected[8:15])
pairs(var_selected[16:25])
```



----- Training and testing the model! ---------------------------

```{r}

# Creating Training and Testing Dataset
set.seed(123)
index <- sample(1:nrow(var_selected), round(nrow(var_selected) * 0.7))
training_df <- var_selected[index, ]
testing_df <- var_selected[-index, ]


```

```{r}
training_df$DIQ010 <- as.integer(training_df$DIQ010)
training_df$DIQ010[training_df$DIQ010 == 1] <- 1
training_df$DIQ010[training_df$DIQ010 == 2] <- 0

null_mod <- glm(DIQ010 ~ 1, data = training_df, family = "binomial")
full_mod <- glm(DIQ010 ~ ., data = training_df, family = "binomial")
best_mod <- step(null_mod, scope = list(lower = null_mod, upper = full_mod),
                       direction="forward", k = 2, trace = 0)
summary(null_mod)
summary(full_mod)
summary(best_mod)
```

LBXGLU - Fasting Glucose (mg/dL)
WTINT2YR - Full Sample 2 Year Interview Weight
PAD675 - Minutes moderate recreational activities
LBXTC - Total Cholesterol (mg/dL)
BMXBMI - Body Mass Index (kg/m**2)
INDFMPIR - Ratio of family income to poverty


```{r}
csq <- 373.73 - 231.79
pchisq(csq, df = 6, lower.tail = F)
```
This means that the model is a good fit.

```{r}
odds <- best_mod$coefficients[1] +
best_mod$coefficients[2]*training_df$LBXGLU +
best_mod$coefficients[3]*training_df$WTINT2YR +
best_mod$coefficients[4]*training_df$PAD675 +  
best_mod$coefficients[5]*training_df$LBXTC +  
best_mod$coefficients[6]*training_df$BMXBMI + 
best_mod$coefficients[7]*training_df$INDFMPIR




library(ggplot2)
ggplot(training_df, aes(x = odds, y = DIQ010)) +
geom_point() +
geom_smooth(method = "glm", method.args = list(family = "binomial"), size = 1.5) +
labs(y = "Probability of diabetes") + theme_bw()
```

Thresholdd of 0.5

```{r}
testing_df$DIQ010 <- as.integer(testing_df$DIQ010)
testing_df$DIQ010[testing_df$DIQ010 == 1] <- 1
testing_df$DIQ010[testing_df$DIQ010 == 2] <- 0

pihat_train <- predict(best_mod,
                       newdata = testing_df,
                       type = "response")
threshold <- 0.5
predicted_category <- factor( ifelse(pihat_train > threshold,
1, 0))
testing_df$DIQ010 <- as.factor(testing_df$DIQ010)

library(caret)

confusionMatrix(data = predicted_category,
reference = (testing_df$DIQ010))
```

88% Accurate!

```{r}
# Taking out insignificant variables. 
sigmod <- glm(DIQ010 ~ LBXGLU + WTINT2YR + PAD675 + LBXTC + BMXBMI, family = "binomial", data = training_df)
summary(sigmod)
```

Significant variables
LBXGLU - Fasting Glucose (mg/dL)
WTINT2YR - Full Sample 2 Year Interview Weight
PAD675 - Minutes moderate recreational activities
LBXTC - Total Cholesterol (mg/dL)
BMXBMI - Body Mass Index (kg/m**2)


```{r}
csq <- 373.73 - 233.95
pchisq(csq, df = 5, lower.tail = F)
```

```{r}
pihat_train <- predict(sigmod,
                       newdata = testing_df[1:24],
                       type = "response")
threshold <- 0.55
predicted_category <- factor( ifelse(pihat_train > threshold,
1, 0))
testing_df$DIQ010 <- as.factor(testing_df$DIQ010)

library(caret)

confusionMatrix(data = predicted_category,
reference = (testing_df$DIQ010))
```
90% accurate!!!


```{r}
pairs(var_selected[c(
"LBXGLU",
"WTINT2YR" ,
"PAD675" ,
"LBXTC" ,
"BMXBMI" 
)])
```

None look colinear which is good.

```{r}
anova(sigmod, best_mod)
```

The anova says that the significant model is better than the stepwise model as well


```{r}
tree_bestmod <- tree(formula = DIQ010 ~ LBXGLU + WTINT2YR + PAD675 + LBXTC + BMXBMI + 
    INDFMPIR, data = training_df)
tree_sigmod <- tree(formula = DIQ010 ~ LBXGLU + WTINT2YR + PAD675 + LBXTC + BMXBMI, data = training_df)
summary(tree_bestmod)
summary(tree_sigmod)
```


Both have a similar residual mean deviance but sigmod is the best. 

Here is the list of different aics. The best one is from the best_mod. 

```{r}
aic_list <- c(556.36, 375.52, 350.02, 375.73, 275.16, 245.79, 245.95)
```




Making a PPROC 

```{r}
library(ROCR)
predictictionVal <- prediction(pihat_train, testing_df$DIQ010)
roc <- performance(predictictionVal,"tpr","fpr")
plot(roc, colorize = T, lwd = 2)
abline(a = 0, b = 1)
```






----CATEGORICAL SUMMARY ANALYSIS--------------
Summary: the categorical data either deletes a ton of diabetes observations or is not significant to the model. 

```{r}
library(tidyverse)
load("C:/Users/laure/git/capstone/chisquareTestFactors.RData")
chisquares %<>% arrange(desc(p_value), desc(rowNum), test_stat)
chisquares
```


```{r}
# Variables that are significant with diabetes. 
sig_factor_vars <- chisquares$name[1:5]
sig_factor_vars
```

```{r}
# Creating dataframe with only significant factor variables. The p-value must be > 0.05  to lack statistically significant evidence 
# that the variables are different from diabetes. 
set.seed(12)
factor_var_df <- selected_df[,which(names(selected_df) %in% sig_factor_vars)]
factor_var_df$DIQ010 <- selected_df$DIQ010
```


Variable Selection with categorical variables:

```{r}
# 405 variables available for modeling
factor_var_df <- subset(factor_var_df, select = -c(DBQ360))
factor_var_df %>% 
  select(one_of(names(factor_var_df))) %>% 
  na.omit() %>% 
  nrow()
factor_var_df <- na.omit(factor_var_df)
summary(factor_var_df)

```


```{r}
index <- sample(1:nrow(factor_var_df), round(nrow(factor_var_df) * 0.7))
cat_train_df <- factor_var_df[index, ]
cat_test_df <- factor_var_df[-index, ]

cat_train_df$DIQ010 <- as.integer(cat_train_df$DIQ010)
cat_train_df$DIQ010[cat_train_df$DIQ010 == 1] <- 1
cat_train_df$DIQ010[cat_train_df$DIQ010 == 2] <- 0

null_mod <- glm(DIQ010 ~ 1, data = cat_train_df, family = "binomial")
full_mod <- glm(DIQ010 ~ ., data = cat_train_df, family = "binomial")
best_mod <- step(null_mod, scope = list(lower = null_mod, upper = full_mod),
                       direction="forward", k = 2, trace = 0)
summary(null_mod)
summary(full_mod)
summary(best_mod)
```

The categorical variables do not look significant at all.

Try to combine with the categorical model:


```{r}

cat_tree <- tree(DIQ010 ~ ., data = cat_test_df)
summary(cat_tree)
```

PFQ041
DBQ9452

I feel like the categorical variables are just not significant.
If they are, they aren't in the NHANES website. Also that AIC value is terrible. 

----------NEURAL NETWORK MODEL-----------------------

```{r}
?nn
```



