---
title: "Quality Analysis"
author: "Lauren Bradley"
date: "11/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Focus Demographics. 

Demographics
Sex/Gender
Age
Race/Ethnicity
Obesity


Chronic diseases
Type 2 diabetes mellitus/Metabolic syndrome
Atherosclerosis
Hypertension
Obesity


Markers
Type 2 diabetes mellitus markers
Glycosylated hemoglobin (Hba1c)
Oral glucose tolerance test
Adiponectin
Glycated albumin
Fructosamine

Metabolic syndrome markers
Waist circumference: >35 inches for females; >40 inches for males
Body mass index: >30 kg/m2 for females and males
Fasting plasma glucose: >100 mg/dL or previously diagnosed type 2 diabetes mellitus
Elevated Triglycerides: >150 mg/dL for females and males
Decreased high density lipoprotein cholesterol (HDL-C): <50 mg/dL for females, <40 mg/dL for males 
Hypertension: Systolic blood pressure: ≥130 mm Hg, Diastolic blood pressure: ≥85 mm Hg

Cardiovascular disease markers
Total cholesterol (TC)
High density lipoprotein cholesterol (HDL-C)
Low density lipoprotein cholesterol (LDL-C)
Triglycerides (TG)
C-reactive Protein (CRP)
Interleukin-6 (IL-6)



Import data

```{r}
load("C:/Users/laure/git/capstone/NHANESCleanFactors.RData")
df_full <- subset(df_full, select = -c(BMIHEAD))

```



Cleaning added according to quality analysis:

```{r}
df_full$DID040[df_full$DID040 == 666 | df_full$DID040 == 999] <- NA
df_full$DIQ010[df_full$DIQ010 == 3 | df_full$DIQ010 == 7 | df_full$DIQ010 == 9] <- 2
df_full$DIQ010[is.na(df_full$DIQ010)] <- 2
```

Determining if people are obese:

```{r}
is_obese_bmi <- c()
for (item in df_full$BMXBMI){
  if (is.na(item)){
    is_obese_bmi <- append(is_obese_bmi, NA)
  }
  else if (item > 30){
    is_obese_bmi <- append(is_obese_bmi, "1")
  }
  else{
    is_obese_bmi <- append(is_obese_bmi, "2")
  }
}

df_full$OBS_BMI <- is_obese_bmi
```

Regression (age vs age when got diabetes)

```{r}
mod <- lm(DID040 ~ RIDAGEYR,data = df_full)
plot(df_full$RIDAGEYR, df_full$DID040)
abline(mod)
```


```{r}
par(mfrow=c(1,3))
plot(mod$fitted.values, mod$residuals, main = "Residual Scatter Plot")
abline(h = 0)

qqnorm(mod$residuals)

hist(mod$residuals)
```





# Try to make a line plot with proportions of whether someone says yes or noo to something

```{r}
# subset dfs by each year
df0910 <- df_full[df_full$Year == "2009-2010",]
df1112 <- df_full[df_full$Year == "2011-2012",]
df1314 <- df_full[df_full$Year == "2013-2014",]
df1516 <- df_full[df_full$Year == "2015-2016",]
df1718 <- df_full[df_full$Year == "2017-2018",]

```

```{r}
dia_prop <- c()
dia_prop <- append(dia_prop,
                  table(df0910$DIQ010)[1]/dim(df0910)[1])
dia_prop <- append(dia_prop,
                  table(df1112$DIQ010)[1]/dim(df1112)[1])
dia_prop <- append(dia_prop, 
                  table(df1314$DIQ010)[1]/dim(df1314)[1])
dia_prop <- append(dia_prop,
                  table(df1516$DIQ010)[1]/dim(df1516)[1])
dia_prop <- append(dia_prop,
                  table(df1718$DIQ010)[1]/dim(df1718)[1])

```


```{r}
library(ggplot2)

prop_df <- data.frame("Year" = c("2009-2010", 
                                 "2011-2012",
                                 "2013-2014",
                                 "2015-2016",
                                 "2017-2018"),
                      "DIA_Prop" = dia_prop)

ggplot(prop_df, aes(x = Year, y = DIA_Prop, group = 1)) + geom_point() + geom_path() + ylab("Diabetes Proportion") + theme_bw() + ggtitle("Diabetes Proportion vs. Year")
```


# Now looking at coronary heart disease

```{r}
heart_prop <- c()
heart_prop <- append(heart_prop,
                  table(df0910$MCQ160B)[1]/dim(df0910)[1])
heart_prop <- append(heart_prop,
                  table(df1112$MCQ160B)[1]/dim(df1112)[1])
heart_prop <- append(heart_prop, 
                  table(df1314$MCQ160B)[1]/dim(df1314)[1])
heart_prop <- append(heart_prop,
                  table(df1516$MCQ160B)[1]/dim(df1516)[1])
heart_prop <- append(heart_prop,
                  table(df1718$MCQ160B)[1]/dim(df1718)[1])

prop_df$HEART_prop <- heart_prop
```

```{r}
ggplot(prop_df, aes(x = Year, y = HEART_prop, group = 1)) + geom_point() + geom_path() + ylab("Congestive Heart Disease Proportion") + theme_bw() + ggtitle("Congestive Heart Disease Proportion vs. Year")

```


```{r}
obs_prop <- c()
obs_prop <- append(obs_prop,
                  table(df0910$BMXBMI)[1]/dim(df0910)[1])
obs_prop <- append(obs_prop,
                  table(df1112$BMXBMI)[1]/dim(df1112)[1])
obs_prop <- append(obs_prop, 
                  table(df1314$BMXBMI)[1]/dim(df1314)[1])
obs_prop <- append(obs_prop,
                  table(df1516$BMXBMI)[1]/dim(df1516)[1])
obs_prop <- append(obs_prop,
                  table(df1718$BMXBMI)[1]/dim(df1718)[1])
prop_df$OBS_prop <- obs_prop
```

```{r}
ggplot(prop_df, aes(x = Year, y = OBS_prop, group = 1)) + geom_point() + geom_path() + ylab("Obesity based on BMI Proportion") + theme_bw() + ggtitle("Obesity vs. Year")
```




# Target Variables
DIQ010

```{r}
one_val <- c()
col_names <- names(df_full)
for (i in 1:length(df_full)){
  if (length(table(df_full[i])) == 1){
    one_val <- col_names[i]
  }
}
```



```{r}
#df_sub <- df_full[,-which(names(df_full) %in% one_val)]

#df_full <- subset(df_full, select = -c(BMIHEAD))
#df_full$DIQ010 <- as.factor(df_full$DIQ010)
null_model <- glm(DIQ010 ~ 1, data = df_full, family = "binomial")
full_model <- glm(DIQ010 ~ ., data = df_full, family = "binomial")
model_step_AIC <- step(null_model,
                       scope = list(lower = null_model, upper = full_model),
                       direction="forward", k = 2, trace = 0)
summary(model_step_AIC)


str(df_full)

```

```{r}
# df_full$DIQ010 <- droplevels(df_full$DIQ010, except = c(1,2))
ggplot(df_full, aes(x=DIQ010, y=RIDAGEYR)) + 
  geom_boxplot() + xlab("Diabetes? 1 - Yes, 2 - No") + ylab("Age") + ggtitle("Age vs. Diabetes")

```

```{r}
ggplot(df_full, aes(x=DIQ010, y=RIDAGEYR)) + 
  geom_boxplot() + xlab("Diabetes? 1 - Yes, 2 - No") + ylab("Age") + ggtitle("Age vs. Diabetes")
```






# Checking for how many null values there are and putting it into a table. 
```{r}
col_names <- names(df_full)
prop_null <- c()
for (i in 1:length(col_names)){
  null_prop <- sum(is.na(df_full[i]))/dim(df_full[i])[1]
  prop_null <- append(prop_null, null_prop)
}

null_df <- data.frame("VariableName" = col_names, "NullProportion" = prop_null)
null_df
```

```{r}
# Available variables
available <- (1 - null_df$NullProportion)*dim(df_full)[1]
null_df$Available <- available
```


```{r}
save(null_df, file = "PropNullDF.RData")
```


# Box plot of null values
```{r}
library(ggplot2)
null_boxplot <- ggplot(null_df, aes(x = "", y = NullProportion)) + 
  geom_boxplot() + xlab("") + ylab("Summary") + ggtitle("Null Proportions of Variables Summary") + theme_bw()
null_boxplot
```

```{r}
num <- null_df$Available[null_df$Available > 3000]
length(num)
```
```{r}
avail_boxplot <- ggplot(null_df, aes(x = "", y = Available)) + 
  geom_boxplot() + xlab("") + ylab("Summary") + ggtitle("Availability of Observations within Variables Summary") + theme_bw()
```

```{r}
avail_boxplot
```


```{r}
ggsave("avail_boxplot.pdf",avail_boxplot)
ggsave("null_boxplot.pdf",null_boxplot)
```


```{r}
number_df <- df_full[,sapply(df_full,is.numeric)]
cor(number_df)

unique(number_df)

```
# -------------------------MODELING--------------


```{r}
table(df_full$DIQ010)
str(df_full$DIQ010)

```
```{r}
col_greater_10000 <- null_df$VariableName[null_df$Available > 10000]

selected_df <- subset(df_full, select = c(col_greater_10000))

```

```{r}
library(tidyverse)
predictorSet <- qs_to_include

include_df <- subset(selected_df, select = c(predictorSet))

#One of the big outliers of 9999 and 7777
include_df$WHD110[include_df$WHD110 == 9999 | include_df$WHD110 == 7777] <- NA

include_df$DIQ010 <- selected_df$DIQ010
include_df$Age <- selected_df$RIDAGEYR
include_df$Gender <- selected_df$RIAGENDR


# Including everything from include_df results in 0
include_df %>% 
  select(one_of(names(predictorSet))) %>% 
  na.omit() %>% 
  nrow()
include_df <- na.omit(include_df)
#selected_df %>% 
#  select(one_of(names(selected_df))) %>% 
#  na.omit() %>% 
#  nrow()
```

Quantitative Variables!


```{r}
mod1 <- glm(DIQ010 ~ LBDHDD
 + LBDHDDSI
 + LBXTR   
 + LBDTRSI 
 + LBDLDL  
 + LBDLDLSI
 + LBXTC
 + LBDTCSI
 + LBXGLU
 + LBDGLUSI
 + CBD121
 + CBD131
 + DBD900
 + DBD905
 + DBD910
 + PAQ625
 + PAD630
 + PAQ670
 + PAD675
 + PAD680
 + WHD010
 + WHD020
 + WHD050
 + WHD110
 + WHD120
 + WHD130 
 + WHD140 
 + WHQ150  
 + BMXWT  
 + BMXHT 
 + BMXBMI
 + BMXLEG 
 + BMXARML 
 + BMXARMC 
 + BMXWAIST
 + BPD035  
 + INDFMPIR
 + WTINT2YR
 + WTMEC2YR + Gender + Age,
 data = include_df, family = "binomial")
summary(mod1)

mod0 <- glm(DIQ010 ~ 1,data = include_df, family = "binomial")

model_step_AIC <- step(mod0,
                       scope = list(lower = mod0, upper = mod1),
                       direction="forward", k = 2, trace = 0)
summary(model_step_AIC)
```



Qualitative Variables:

```{r}
#str(selected_df)
cat_df <- selected_df[,sapply(selected_df,is.factor)]
dim(cat_df)

```
154 variables 

```{r}
library(tidyverse)
cat_df %>% 
  select(one_of(names(cat_df))) %>% 
  na.omit() %>% 
  nrow()
#cat_df <- na.omit(cat_df)
```

```{r}
cat_null <- null_df[is.element(null_df$VariableName,c(names(cat_df))),]
```

```{r}
barplot(cat_null$Available)
```

```{r}
cat_greater_30000 <- cat_null$VariableName[cat_null$Available > 10000]

cat_selected <- subset(cat_df, select = c(cat_greater_30000))

cat_selected %>% 
  select(one_of(names(cat_selected))) %>% 
  na.omit() %>% 
  nrow()

cat_selected <- na.omit(cat_selected)

summary(cat_selected)
remove_these <- c("AIALANGA",
                  "BMDSTATS",
                  "CBQ695",
                  "CBQ695.1",
                  "CBQ698",
                  "CBQ698.1",
                  "CBQ738",
                  "CBQ738.1")

# ISSUE: the DIQ010 does not have literally any yes responses, not good..

```



```{r}
full_cat_mod <- glm(DIQ010 ~ .,
 data = cat_selected, family = "binomial")
summary(full_cat_mod)

mod0 <- glm(DIQ010 ~ 1,data = include_df, family = "binomial")

model_step_AIC <- step(mod0,
                       scope = list(lower = mod0, upper = mod1),
                       direction="forward", k = 2, trace = 0)
summary(model_step_AIC)
```
```{r}
save(selected_df, file = "selected_df.RData")
```





-------------------Jump to going back to quantitative values-------------------


Mod2 is a similar model to model_step_aic but without the insignificant variables

```{r}
mod2 <- glm(formula = DIQ010 ~ LBDGLUSI + LBDTCSI + PAD675 + 
    WTMEC2YR + BMXARML + WHD140 + LBDHDDSI + WHD110, family = "binomial", data = include_df)
summary(mod2)
```
```{r}
anova(mod2, model_step_AIC)
```

Mod2 is a better model according to the anova. Lets see if any are multicolinear:

```{r}
mod3 <- glm(formula = DIQ010 ~ LBDGLUSI + LBDTCSI + PAD675 + BMXARML + WHD110 + LBDHDDSI, family = "binomial", data = include_df)
summary(mod3)
```
```{r}
anova(mod3, model_step_AIC)
```


```{r}
anova(mod3, mod2)
```



```{r}
mod_vars <- subset(include_df, select = c(LBDGLUSI , LBDTCSI , PAD675 , BMXARML , 
    WHD110 ,LBDHDDSI))

pairs(mod_vars)
```

BMXARML and WHD140 are colinear. Need to modify mod3 because of this. 

```{r}
mod4 <- glm(formula = DIQ010 ~ LBDGLUSI + LBDTCSI + PAD675 + BMXARML*WHD110 + LBDHDDSI, family = "binomial", data = include_df)
summary(mod4)
```

```{r}
anova(mod3, mod4)
```

```{r}
summary(mod3)
```

Since BMXARML is more significant than WHD110, we are going to keep that variable and remove whd110

```{r}
mod5 <- glm(DIQ010 ~ LBDGLUSI + LBDTCSI + PAD675 + BMXARML + LBDHDDSI, family = "binomial", data = include_df)
summary(mod5)
```

LBDGLUSI - Fasting Glucose (mmol/L)
LBDTCSI - Total Cholesterol (mmol/L)
PAD675 - Minutes moderate recreational activities
BMXARML - Upper Arm Length (cm)
LBDHDDSI - Direct HDL-Cholesterol (mmol/L)




```{r}
library(tree)
tree1 <- tree(DIQ010 ~ LBDHDD
 + LBDHDDSI
 + LBXTR   
 + LBDTRSI 
 + LBDLDL  
 + LBDLDLSI
 + LBXTC
 + LBDTCSI
 + LBXGLU
 + LBDGLUSI
 + CBD121
 + CBD131
 + DBD900
 + DBD905
 + DBD910
 + PAQ625
 + PAD630
 + PAQ670
 + PAD675
 + PAD680
 + WHD010
 + WHD020
 + WHD050
 + WHD110
 + WHD120
 + WHD130 
 + WHD140 
 + WHQ150  
 + BMXWT  
 + BMXHT 
 + BMXBMI
 + BMXLEG 
 + BMXARML 
 + BMXARMC 
 + BMXWAIST
 + BPD035  
 + INDFMPIR
 + WTINT2YR
 + WTMEC2YR + Gender + Age, data = include_df)

summary(tree1)
```

Tree 1 Variables Used:

LBXGLU - Fasting Glucose (mg/dL)
BMXWT - Weight (kg)
WTINT2YR - Full Sample 2 Year Interview Weight
INDFMPIR - Ratio of family income to poverty
WTMEC2YR - Full Sample 2 Year MEC Exam Weight
LBXTC - Total Cholesterol (mg/dL)
CBD131 - Money spent on carryout/delivered foods
BPD035 - Age told had hypertension
LBDHDD - Direct HDL Cholesterol (mg/dL)
LBXTR - Triglyceride (mg/dL)
WHD020 - Current self-reported weight (pounds)
WHD010 - Current self-reported height (inches)
LBDLDL - LDL-Cholesterol, Friedewald (mg/dL)
PAQ625 - Number of days moderate work
WHD130 - Self-reported height -age 25 (inches)




```{r}
plot(tree1)
text(tree1, cex = 0.7)
```


```{r}
tree_df <- subset(include_df, select = c("LBXGLU" ,  "BMXWT"  ,  "WTINT2YR", "INDFMPIR" ,"WTMEC2YR", "LBXTC"  ,  "CBD131" ,  "BPD035" ,  "LBDHDD"   ,"LBXTR",    "WHD020" ,  "WHD010" ,  "LBDLDL" ,  "PAQ625" , "WHD130"))
pairs(tree_df[1:5])
pairs(tree_df[5:10])
pairs(tree_df[10:15])
```

Vars with outliers: bpd035, cbd131, whd130. wtint2yr and wtmec2yr are too correlated. 

```{r}
table(include_df$WHD130)
include_df <- include_df[include_df$BPD035 != 999]
include_df <- include_df[include_df$CBD131 != 777777 | include_df$CBD131 != 999999]
include_df <- include_df[include_df$WHD130 != 9999]
```





```{r}
library(randomForest)

include_df$DIQ010 <- droplevels(include_df$DIQ010)


rforest <- randomForest(formula = DIQ010 ~ LBDHDD
 + LBDHDDSI
 + LBXTR   
 + LBDTRSI 
 + LBDLDL  
 + LBDLDLSI
 + LBXTC
 + LBDTCSI
 + LBXGLU
 + LBDGLUSI
 + CBD121
 + CBD131
 + DBD900
 + DBD905
 + DBD910
 + PAQ625
 + PAD630
 + PAQ670
 + PAD675
 + PAD680
 + WHD010
 + WHD020
 + WHD050
 + WHD110
 + WHD120
 + WHD130 
 + WHD140 
 + WHQ150  
 + BMXWT  
 + BMXHT 
 + BMXBMI
 + BMXLEG 
 + BMXARML 
 + BMXARMC 
 + BMXWAIST
 + BPD035  
 + INDFMPIR
 + WTINT2YR
 + WTMEC2YR + Gender + Age, data = include_df, ntree = 250, nodesize = 25, importance = TRUE)

varImpPlot(rforest, type = 1, main = "")
```

Top Ten Variables

LBDGLUSI - Fasting Glucose (mmol/L)
-- LBXGLU - Fasting Glucose (mg/dL)
LBDLDLSI - LDL-Cholesterol, Friedewald (mmol/L)
LBDTCSI - Total Cholesterol (mmol/L)
WHD110 - Self-reported weight-10 yrs ago (pounds)
--LBXTC - Total Cholesterol (mg/dL)
--WHD050 - Self-reported weight-1 yr ago (pounds)
--WHD140 - Self-reported greatest weight (pounds)
--LBDLdL - LDL-Cholesterol, Friedewald (mg/dL)
BMXWaist - Waist Circumference (cm)
BMXARML - Upper Arm Length (cm)





------ MAIN VARIABLES I AM INTERESTED IN NOW -----------------------------

LBDGLUSI - Fasting Glucose (mmol/L)
LBDLDLSI - LDL-Cholesterol, Friedewald (mmol/L)
LBDTCSI - Total Cholesterol (mmol/L)
WHD110 - Self-reported weight-10 yrs ago (pounds) #
BMXWaist - Waist Circumference (cm)               # Potentially Co linear
BMXARML - Upper Arm Length (cm)                   #
BMXWT - Weight (kg)
INDFMPIR - Ratio of family income to poverty
CBD131 - Money spent on carryout/delivered foods
BPD035 - Age told had hypertension
PAQ625 - Number of days moderate work
WHD130 - Self-reported height -age 25 (inches)
PAD675 - Minutes moderate recreational activities
LBDHDDSI - Direct HDL-Cholesterol (mmol/L)



```{r}
# Cleaning variables
selected_df$BPD035[selected_df$BPD035 == 999] <- NA
selected_df$CBD131[selected_df$CBD131 == 777777 | selected_df$CBD131 == 999999] <- NA
selected_df$WHD130[selected_df$WHD130 == 9999] <- NA
selected_df$WHD110[selected_df$WHD110 == 9999] <- NA
selected_df$PAD675[selected_df$PAD675 == 9999] <- NA



var_selected <- subset(selected_df, select = c(
  LBDGLUSI,
  LBDLDLSI,
  LBDTCSI,
  WHD110,
  BMXWAIST,
  BMXARML,
  BMXWT,
  INDFMPIR,
  CBD131,
  BPD035,
  PAQ625,
  WHD130,
  PAD675,
  LBDHDDSI,
  DIQ010
))


```



```{r}
var_selected %>% 
  select(one_of(names(var_selected))) %>% 
  na.omit() %>% 
  nrow()
var_selected <- na.omit(var_selected)

var_selected$DIQ010 <- droplevels(var_selected$DIQ010)
```



```{r}
null_mod <- glm(DIQ010 ~ 1, data = var_selected, family = "binomial")
full_mod <- glm(DIQ010 ~ ., data = var_selected, family = "binomial")
best_mod <- step(null_mod, scope = list(lower = null_mod, upper = full_mod),
                       direction="forward", k = 2, trace = 0)
summary(best_mod)
```


```{r}
tree2 <- tree(DIQ010 ~ LBDGLUSI + LBDTCSI + PAD675 + BMXWAIST + 
    WHD130 + WHD110 + INDFMPIR, data = var_selected)
summary(tree2)
```

```{r}
plot(tree2)
text(tree2, cex = 0.7)
```



```{r}
rforest2 <- randomForest(DIQ010 ~ ., data = var_selected, ntree = 250, nodesize = 25, importance = TRUE)
varImpPlot(rforest2, type = 1, main = "")
```

```{r}
pairs(var_selected[1:7])
pairs(var_selected[8:15])
```


----- Training and testing the model! ---------------------------

```{r}

# Creating Training and Testing Dataset
set.seed(1234)
index <- sample(1:nrow(var_selected), round(nrow(var_selected) * 0.7))
training_df <- var_selected[index, ]
testing_df <- var_selected[-index, ]


```

```{r}
training_df$DIQ010 <- as.integer(training_df$DIQ010)
training_df$DIQ010[training_df$DIQ010 == 1] <- 1
training_df$DIQ010[training_df$DIQ010 == 2] <- 0

null_mod <- glm(DIQ010 ~ 1, data = training_df, family = "binomial")
full_mod <- glm(DIQ010 ~ ., data = training_df, family = "binomial")
best_mod <- step(null_mod, scope = list(lower = null_mod, upper = full_mod),
                       direction="forward", k = 2, trace = 0)
summary(best_mod)
```

Same model as the one created without the train!


```{r}
csq <- 299.14 - 153.87
pchisq(csq, df = 7, lower.tail = F)
```
This means that the model is a good fit.

```{r}
odds <- best_mod$coefficients[1] +
best_mod$coefficients[2]*training_df$LBDGLUSI +
best_mod$coefficients[3]*training_df$LBDTCSI +
best_mod$coefficients[4]*training_df$PAD675 + 
best_mod$coefficients[5]*training_df$INDFMPIR + 
best_mod$coefficients[6]*training_df$BMXWAIST +
best_mod$coefficients[7]*training_df$WHD130 +
best_mod$coefficients[8]*training_df$WHD110




library(ggplot2)
ggplot(training_df, aes(x = odds, y = DIQ010)) +
geom_point() +
geom_smooth(method = "glm", method.args = list(family = "binomial"), size = 1.5) +
labs(y = "Probability of Strong Agreement") + theme_bw()
```

Thresholdd of 0.5

```{r}
testing_df$DIQ010 <- as.integer(testing_df$DIQ010)
testing_df$DIQ010[testing_df$DIQ010 == 1] <- 1
testing_df$DIQ010[testing_df$DIQ010 == 2] <- 0

pihat_train <- predict(best_mod,
                       newdata = testing_df,
                       type = "response")
threshold <- 0.50
predicted_category <- factor( ifelse(pihat_train > threshold,
1, 0))
testing_df$DIQ010 <- as.factor(testing_df$DIQ010)

library(caret)

confusionMatrix(data = predicted_category,
reference = (testing_df$DIQ010))
```

79% Accurate. 

```{r}
# Taking out insignificant variables. 
sigmod <- glm(formula = DIQ010 ~ LBDGLUSI + LBDTCSI + PAD675 + INDFMPIR, family = "binomial", data = training_df)
summary(sigmod)
```

LBDGLUSI - Fasting Glucose (mmol/L)
LBDTCSI - Total Cholesterol (mmol/L)
PAD675 - Minutes moderate recreational activities
INDFMPIR - Ratio of family income to poverty

```{r}
csq <- 299.14 - 164.85
pchisq(csq, df = 4, lower.tail = F)
```

```{r}
pihat_train <- predict(sigmod,
                       newdata = testing_df,
                       type = "response")
threshold <- 0.50
predicted_category <- factor( ifelse(pihat_train > threshold,
1, 0))
testing_df$DIQ010 <- as.factor(testing_df$DIQ010)

library(caret)

confusionMatrix(data = predicted_category,
reference = (testing_df$DIQ010))
```
80% accurate. So more accurate than the step wise model. 


```{r}
pairs(var_selected[c("LBDGLUSI", "LBDTCSI", "PAD675", "INDFMPIR")])
```

None look colinear which is good.


```{r}
original_select_df <- selected_df
```


--------------------Now done with outliers, new analysis!-------------------------------------------

```{r}
library(tidyverse)
predictorSet <- qs_to_include

include_df <- subset(selected_df, select = c(predictorSet))

#One of the big outliers of 9999 and 7777
include_df$WHD110[include_df$WHD110 == 9999 | include_df$WHD110 == 7777] <- NA
include_df$WHD120[include_df$WHD120 == 9999] <- NA

include_df$DIQ010 <- selected_df$DIQ010
include_df$Age <- selected_df$RIDAGEYR
include_df$Gender <- selected_df$RIAGENDR


# Including everything from include_df results in 0
include_df %>% 
  select(one_of(names(predictorSet))) %>% 
  na.omit() %>% 
  nrow()
include_df <- na.omit(include_df)
#selected_df %>% 
#  select(one_of(names(selected_df))) %>% 
#  na.omit() %>% 
#  nrow()
```

Quantitative Variables!


```{r}
mod1 <- glm(DIQ010 ~ LBDHDD
 + LBDHDDSI
 + LBXTR   
 + LBDTRSI 
 + LBDLDL  
 + LBDLDLSI
 + LBXTC
 + LBDTCSI
 + LBXGLU
 + LBDGLUSI
 + CBD121
 + CBD131
 + DBD900
 + DBD905
 + DBD910
 + PAQ625
 + PAD630
 + PAQ670
 + PAD675
 + PAD680
 + WHD010
 + WHD020
 + WHD050
 + WHD110
 + WHD120
 + WHD130 
 + WHD140 
 + WHQ150  
 + BMXWT  
 + BMXHT 
 + BMXBMI
 + BMXLEG 
 + BMXARML 
 + BMXARMC 
 + BMXWAIST
 + BPD035  
 + INDFMPIR
 + WTINT2YR
 + WTMEC2YR + Gender + Age,
 data = include_df, family = "binomial")
summary(mod1)

mod0 <- glm(DIQ010 ~ 1,data = include_df, family = "binomial")

model_step_AIC <- step(mod0,
                       scope = list(lower = mod0, upper = mod1),
                       direction="forward", k = 2, trace = 0)
summary(model_step_AIC)
```


LBXGLU - Fasting Glucose (mg/dL)
LBXTC - Total Cholesterol (mg/dL)
PAD675 - Minutes moderate recreational activities
BMXBMI - Body Mass Index (kg/m**2)
WTMEC2YR - Full Sample 2 Year MEC Exam Weight
PAQ670 - Days moderate recreational activities
LBDHDDSI - Direct HDL-Cholesterol (mmol/L)
BMXARML - Upper Arm Length (cm)
WHD110 - Self-reported weight-10 yrs ago (pounds)
CBD121 - Money spent on eating out



-------------------Jump to going back to quantitative values-------------------


Mod2 is a similar model to model_step_aic but without the insignificant variables

```{r}
mod2 <- glm(formula = DIQ010 ~ LBXGLU + LBXTC + PAD675 + BMXBMI + WTMEC2YR + 
    PAQ670 + LBDHDDSI + BMXARML + WHD110, family = "binomial", 
    data = include_df)
summary(mod2)
```


```{r}
anova(mod2, model_step_AIC)
```


```{r}
mod3 <- glm(formula = DIQ010 ~ LBXGLU + LBXTC + PAD675 + WTMEC2YR + 
    PAQ670 + LBDHDDSI + BMXARML + WHD110, family = "binomial", 
    data = include_df)
summary(mod3)
```

```{r}
anova(mod3, model_step_AIC)
```


```{r}
anova(mod3, mod2)
```



```{r}
mod_vars <- subset(include_df, select = c(
  LBXGLU ,
LBXTC  ,
PAD675 ,
WTMEC2YR , 
PAQ670  ,
LBDHDDSI  ,  
BMXARML  ,
WHD110
))

pairs(mod_vars)
```

BMXARML and WHD110 are colinear. Need to modify mod3 because of this, and removing uhhh IDK

```{r}
library(tree)
tree0 <- tree(DIQ010 ~ LBXGLU + LBXTC + PAD675 + WTMEC2YR + 
    PAQ670 + LBDHDDSI + BMXARML + WHD110, 
    data = include_df)
summary(tree0)
```


```{r}
mod4 <- glm(formula = DIQ010 ~ LBXGLU + LBXTC + PAD675 + WTMEC2YR + 
    PAQ670 + LBDHDDSI + WHD110, family = "binomial", 
    data = include_df)
summary(mod4)
```

```{r}
anova(mod4, mod3)
```


```{r}
mod5 <- glm(formula = DIQ010 ~ LBXGLU + LBXTC + PAD675 + WTMEC2YR + 
    PAQ670 + WHD110, family = "binomial", 
    data = include_df)
summary(mod5)
```

```{r}
anova(mod5, mod4)
```

Best Potential Model:

LBXGLU - Fasting Glucose (mg/dL)
LBXTC - Total Cholesterol (mg/dL)
PAD675 - Minutes moderate recreational activities
WTMEC2YR - Full Sample 2 Year MEC Exam Weight
PAQ670 - Days moderate recreational activities
WHD110 - Self-reported weight-10 yrs ago (pounds)

```{r}
tree1 <- tree(DIQ010 ~ LBDHDD
 + LBDHDDSI
 + LBXTR   
 + LBDTRSI 
 + LBDLDL  
 + LBDLDLSI
 + LBXTC
 + LBDTCSI
 + LBXGLU
 + LBDGLUSI
 + CBD121
 + CBD131
 + DBD900
 + DBD905
 + DBD910
 + PAQ625
 + PAD630
 + PAQ670
 + PAD675
 + PAD680
 + WHD010
 + WHD020
 + WHD050
 + WHD110
 + WHD120
 + WHD130 
 + WHD140 
 + WHQ150  
 + BMXWT  
 + BMXHT 
 + BMXBMI
 + BMXLEG 
 + BMXARML 
 + BMXARMC 
 + BMXWAIST
 + BPD035  
 + INDFMPIR
 + WTINT2YR
 + WTMEC2YR + Gender + Age, data = include_df)

summary(tree1)
```

Tree 1 Variables Used:

LBXGLU - Fasting Glucose (mg/dL)
BMXWT - Weight (kg)
WTINT2YR - Full Sample 2 Year Interview Weight
PAQ625 - Number of days moderate work
WHD120 - Self-reported weight-age 25 (pounds)
BPD035 - Age told had hypertension
INDFMPIR - Ratio of family income to poverty
LBDHDD - Direct HDL Cholesterol (mg/dL)
LBXTR - Triglyceride (mg/dL)
WHD020 - Current self-reported weight (pounds)
WHD010 - Current self-reported height (inches)
Age
LBXTC - Total Cholesterol (mg/dL)
DBD910 - # of frozen meals/pizza in past 30 days



```{r}
plot(tree1)
text(tree1, cex = 0.7)
```


```{r}
tree_df <- subset(include_df, select = c(
  "LBXGLU" ,  "BMXWT"   , "WTINT2YR", "PAQ625"  , "WHD120",   "BPD035"  , "INDFMPIR", "LBDHDD"  , "LBXTR" ,   "WHD020"  , "WHD010"  , "Age",     
 "LBXTC"  ,  "DBD910"  
))
pairs(tree_df[1:5])
pairs(tree_df[5:10])
pairs(tree_df[10:14])
```

Vars with outliers: bpd035, cbd131, whd130. wtint2yr and wtmec2yr are too correlated. 

```{r}
table(include_df$WHD130)
include_df <- include_df[include_df$BPD035 != 999]
include_df <- include_df[include_df$CBD131 != 777777 | include_df$CBD131 != 999999]
include_df <- include_df[include_df$WHD130 != 9999]
```





```{r}
library(randomForest)

include_df$DIQ010 <- droplevels(include_df$DIQ010)


rforest <- randomForest(formula = DIQ010 ~ LBDHDD
 + LBDHDDSI
 + LBXTR   
 + LBDTRSI 
 + LBDLDL  
 + LBDLDLSI
 + LBXTC
 + LBDTCSI
 + LBXGLU
 + LBDGLUSI
 + CBD121
 + CBD131
 + DBD900
 + DBD905
 + DBD910
 + PAQ625
 + PAD630
 + PAQ670
 + PAD675
 + PAD680
 + WHD010
 + WHD020
 + WHD050
 + WHD110
 + WHD120
 + WHD130 
 + WHD140 
 + WHQ150  
 + BMXWT  
 + BMXHT 
 + BMXBMI
 + BMXLEG 
 + BMXARML 
 + BMXARMC 
 + BMXWAIST
 + BPD035  
 + INDFMPIR
 + WTINT2YR
 + WTMEC2YR + Gender + Age, data = include_df, ntree = 250, nodesize = 25, importance = TRUE)

varImpPlot(rforest, type = 1, main = "")
```

Top Ten Variables

LBDGLUSI - Fasting Glucose (mmol/L)
-- LBXGLU - Fasting Glucose (mg/dL)
LBDLDLSI - LDL-Cholesterol, Friedewald (mmol/L)
LBDTCSI - Total Cholesterol (mmol/L)
WHD110 - Self-reported weight-10 yrs ago (pounds)
--LBXTC - Total Cholesterol (mg/dL)
--WHD050 - Self-reported weight-1 yr ago (pounds)
--WHD140 - Self-reported greatest weight (pounds)
--LBDLdL - LDL-Cholesterol, Friedewald (mg/dL)
BMXWaist - Waist Circumference (cm)
BMXARML - Upper Arm Length (cm)





------ MAIN VARIABLES I AM INTERESTED IN NOW -----------------------------

LBXGLU - Fasting Glucose (mg/dL)
BMXWT - Weight (kg)
WTINT2YR - Full Sample 2 Year Interview Weight
PAQ625 - Number of days moderate work
WHD120 - Self-reported weight-age 25 (pounds)
BPD035 - Age told had hypertension
INDFMPIR - Ratio of family income to poverty
LBDHDD - Direct HDL Cholesterol (mg/dL)
LBXTR - Triglyceride (mg/dL)
WHD020 - Current self-reported weight (pounds)
WHD010 - Current self-reported height (inches)
Age
LBXTC - Total Cholesterol (mg/dL)
DBD910 - # of frozen meals/pizza in past 30 days
PAD675 - Minutes moderate recreational activities
WTMEC2YR - Full Sample 2 Year MEC Exam Weight
PAQ670 - Days moderate recreational activities
WHD110 - Self-reported weight-10 yrs ago (pounds)
BMXBMI - Body Mass Index (kg/m**2)
LBDHDDSI - Direct HDL-Cholesterol (mmol/L)
BMXARML - Upper Arm Length (cm)
CBD121 - Money spent on eating out




```{r}
# Cleaning variables
selected_df$WHD120[selected_df$WHD120 == 9999] <- NA
selected_df$WHD110[selected_df$WHD110 == 9999 | selected_df$WHD110 == 7777] <- NA
selected_df$WHD020[selected_df$WHD020 == 9999 | selected_df$WHD020 == 7777] <- NA
selected_df$WHD010[selected_df$WHD010 == 9999 | selected_df$WHD010 == 7777] <- NA
selected_df$CBD121[selected_df$CBD121 == 999999 | selected_df$CBD121 == 777777] <- NA

var_selected <- subset(selected_df, select = c(
  LBXGLU ,
BMXWT ,
WTINT2YR ,
PAQ625 ,
WHD120 ,
BPD035 ,
INDFMPIR ,
LBDHDD ,
LBXTR ,
WHD020 ,
WHD010 ,
RIDAGEYR, 
LBXTC ,
DBD910 ,
PAD675 ,
WTMEC2YR ,
PAQ670 ,
WHD110 ,
BMXBMI ,
LBDHDDSI ,
BMXARML ,
CBD121 ,
  DIQ010
))

head(var_selected)
```



```{r}
var_selected %>% 
  select(one_of(names(var_selected))) %>% 
  na.omit() %>% 
  nrow()
var_selected <- na.omit(var_selected)

var_selected$DIQ010 <- droplevels(var_selected$DIQ010)
```



```{r}
null_mod <- glm(DIQ010 ~ 1, data = var_selected, family = "binomial")
full_mod <- glm(DIQ010 ~ ., data = var_selected, family = "binomial")
best_mod <- step(null_mod, scope = list(lower = null_mod, upper = full_mod),
                       direction="forward", k = 2, trace = 0)
summary(best_mod)
```


```{r}
tree2 <- tree(DIQ010 ~ LBXGLU + WTINT2YR + PAD675 + LBXTC + BMXBMI + 
    RIDAGEYR, data = var_selected)
summary(tree2)
```

```{r}
plot(tree2)
text(tree2, cex = 0.7)
```



```{r}
rforest2 <- randomForest(DIQ010 ~ ., data = var_selected, ntree = 250, nodesize = 25, importance = TRUE)
varImpPlot(rforest2, type = 1, main = "")
```

```{r}
pairs(var_selected[1:7])
pairs(var_selected[8:15])
pairs(var_selected[16:23])
```


----- Training and testing the model! ---------------------------

```{r}

# Creating Training and Testing Dataset
set.seed(1234)
index <- sample(1:nrow(var_selected), round(nrow(var_selected) * 0.7))
training_df <- var_selected[index, ]
testing_df <- var_selected[-index, ]


```

```{r}
training_df$DIQ010 <- as.integer(training_df$DIQ010)
training_df$DIQ010[training_df$DIQ010 == 1] <- 1
training_df$DIQ010[training_df$DIQ010 == 2] <- 0

null_mod <- glm(DIQ010 ~ 1, data = training_df, family = "binomial")
full_mod <- glm(DIQ010 ~ ., data = training_df, family = "binomial")
best_mod <- step(null_mod, scope = list(lower = null_mod, upper = full_mod),
                       direction="forward", k = 2, trace = 0)
summary(best_mod)
```

LBXGLU - Fasting Glucose (mg/dL)
WTINT2YR - Full Sample 2 Year Interview Weight
BMXBMI - Body Mass Index (kg/m**2)
LBXTC - Total Cholesterol (mg/dL)
PAD675 - Minutes moderate recreational activities
WHD010 - Current self-reported height (inches)
INDFMPIR - Ratio of family income to poverty


```{r}
csq <- 299.14 - 153.87
pchisq(csq, df = 7, lower.tail = F)
```
This means that the model is a good fit.

```{r}
odds <- best_mod$coefficients[1] +
best_mod$coefficients[2]*training_df$LBXGLU +
best_mod$coefficients[3]*training_df$WTINT2YR +
best_mod$coefficients[4]*training_df$BMXBMI + 
best_mod$coefficients[5]*training_df$LBXTC + 
best_mod$coefficients[6]*training_df$PAD675 +
best_mod$coefficients[7]*training_df$WHD010 +
best_mod$coefficients[8]*training_df$INDFMPIR




library(ggplot2)
ggplot(training_df, aes(x = odds, y = DIQ010)) +
geom_point() +
geom_smooth(method = "glm", method.args = list(family = "binomial"), size = 1.5) +
labs(y = "Probability of diabetes") + theme_bw()
```

Thresholdd of 0.5

```{r}
testing_df$DIQ010 <- as.integer(testing_df$DIQ010)
testing_df$DIQ010[testing_df$DIQ010 == 1] <- 1
testing_df$DIQ010[testing_df$DIQ010 == 2] <- 0

pihat_train <- predict(best_mod,
                       newdata = testing_df,
                       type = "response")
threshold <- 0.5
predicted_category <- factor( ifelse(pihat_train > threshold,
1, 0))
testing_df$DIQ010 <- as.factor(testing_df$DIQ010)

library(caret)

confusionMatrix(data = predicted_category,
reference = (testing_df$DIQ010))
```

82% Accurate. 

```{r}
# Taking out insignificant variables. 
sigmod <- glm(formula = DIQ010 ~ LBXGLU + BMXBMI + LBXTC + PAD675 + 
    WHD010 + INDFMPIR, family = "binomial", data = training_df)
summary(sigmod)
```

Significant variables
LBXGLU - Fasting Glucose (mg/dL)
BMXBMI - Body Mass Index (kg/m**2)
LBXTC - Total Cholesterol (mg/dL)
PAD675 - Minutes moderate recreational activities
WHD010 - Current self-reported height (inches)
INDFMPIR - Ratio of family income to poverty

```{r}
csq <- 299.14 - 164.85
pchisq(csq, df = 4, lower.tail = F)
```

```{r}
pihat_train <- predict(sigmod,
                       newdata = testing_df,
                       type = "response")
threshold <- 0.55
predicted_category <- factor( ifelse(pihat_train > threshold,
1, 0))
testing_df$DIQ010 <- as.factor(testing_df$DIQ010)

library(caret)

confusionMatrix(data = predicted_category,
reference = (testing_df$DIQ010))
```
81% accurate. So less accurate which is interesting


```{r}
pairs(var_selected[c("LBXGLU",
"BMXBMI",
"LBXTC",
"PAD675",
"WHD010",
"INDFMPIR",
"WTINT2YR")])
```

None look colinear which is good, however the step wise model looked better than the significant model

```{r}
anova(sigmod, best_mod)
```




```{r}
tree_bestmod <- tree(DIQ010 ~ LBXGLU + WTINT2YR + BMXBMI + LBXTC + PAD675 + WHD010 + 
    INDFMPIR, data = training_df)
tree_sigmod <- tree(DIQ010 ~ LBXGLU + BMXBMI + LBXTC + PAD675 + WHD010 + 
    INDFMPIR, data = training_df)
summary(tree_bestmod)
summary(tree_sigmod)
```

This one also has tree_bestmod as a better model. 

- proc package
- ROC Curve: determines the sensitivity
- tbRoc = pROC::roc(response = tbCoords$improved, predictor = tbRf$votes[,1])
- nn(formula, data = data)




---------------------Look at different variables? Heart disease?----------------















